\mythesistitleNOTSYNPOSIS \tnote[fyi]{Synopsis title is ``Generalization of Consecutive Ones Property of Binary Matrices''}

### 2012-01-31 src/030chap3.tex
- Figure 3.1 for hypergraphs added. Preliminaries section complete. - Characterization of FTPL section moreorless done except section for worked out example -- changed h.isomorphism induced tpl notation to \cp_{\phi} since \cl is confusing because it is the input in most places - Filled up some definitions in Chap 1 - Added asteroidal triples def in chapter 2

\subsection{Set systems and hypergraphs}
%\begin{definition}[Set systems and hypergraphs]

  The set $\F \subseteq (2^{U} \setminus \emptyset)$ is a {\em set
    system} of a universe $U$ with $|U| = n$.  The {\em support} of a
  set system $\F$ denoted by $\supt{\cF}$ is the union of all the sets
  in $\F$; $supp(\F) = \bigcup_{S \in \F}S$. For the purposes of this
  paper, a set system is required to ``cover'' the universe; $
  supp(\cF) = U$.
  \par
%  \noindent
  A set system $\cF$ can also be visualized as a {\em hypergraph}
  whose vertex set is $supp(\cF)$ and hyperedges are the sets in
  $\cF$. This is a known representation for
  interval systems in literature \cite{bls99,kklv10}.  We extend this
  definition here to path systems. Due to the equivalence of set
  system and hypergraph in the scope of this paper, we drop the
  subscript $_H$ in the notation and refer to both the structures by
  $\cF$.
  \par
%  \noindent
  The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph
  $\cF$ is a graph such that its vertex set has a bijection to $\cF$
  and there exists an edge between two vertices \iff their
  corresponding hyperedges have a non-empty intersection \cite{mcg04}.
  \par
%  \noindent 
  Two hypergraphs $\cF'$, $\cF''$ are said to be {\em isomorphic} to
  each other, denoted by $\cF' \cong \cF''$, \iff there exists a
  bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such that for
  all sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in $\cF'$ \iff
  $B$ is a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x \in A\}$
  \cite{kklv10}, written as $B=\phi(A)$.  This is called {\em
    hypergraph isomorphism}.  
%\end{definition}

\subsection{Path Hypergraph from a Tree}
%\begin{definition}[Path Hypergraph from a Tree]
  The graph $T$ represents a {\em target tree} with same number of
  vertices as elements in $U$; $|V(T)|=|U|=n$.  A {\em path system}\,
  $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P \mid P
  \subseteq V, \text{ } T[P] \text{ is a path} \}$.  This generalizes
  the fact, from the literature \cite{bls99,kklv10}, that intervals
  can be viewed as sub-paths of a path.

%  \noindent
  If the intersection graphs of $\cF$ and $\cP$ (a path system) are
  isomorphic, $\bI(\cF) \cong \bI(\cP)$, then the associated bijection
  $\cl: \cF \rightarrow \cP$ due to this isomorphism is called a {\em
    path labeling} of the hypergraph $\cF$.\footnote{Note that there
    are two kinds of isomorphisms here.  One is the isomorphism of
    intersection graphs on $\cF$ and $\cP$, i.e. $\bI(\cF)$ and
    $\bI(\cP)$ respectively. Second is the isomorphism between the
    hypergraphs $\cF$ and $\cP$.}

% The subset of the path labeling that
%   maps elements of the universe to paths that contain only a leaf the
%   target tree is  known as a {\em leaf labeling}.
%   \ie if the target tree has the set of $k$ leaves $L = \{v_i, i \in
%   [k]\}$, then 
% $\cl': \rightarrow U' \rightarrow L$ is a leaf labeling where $U'
% \subseteq U$, $|U'| = k$.

%   $\cL_0 = \{ (x_i,v_i) \mid x_i \in
%   U, v_i \in T, x_i \ne x_j, v_i \ne v_j, i \ne j, i,j \in [k] \}$
%   where $k$ is the number of leaves in $T$

%  \noindent
  \add{To illustrate further, let $\cg: V(\cF) \rightarrow V(\cP)$ be
    the above mentioned isomorphism where $V(\cF)$ and $V(\cP)$ are
    the vertex sets that represent the hyperedges for each hypergraph
    respectively, $V(\cF) = \{ v_S \mid S \in \cF\}$ and $V(\cP) = \{
    v_P \mid P \in \cP\}$. Then the path labeling $\cl$ is defined as
    follows: $\cl(S_1) = P_1$ iff $\cg (v_{S_1}) = v_{P_1}$.  Just to
    emphasize, for a path labeling $\cl$ of $\cF$ with $\cP$ as the
    path system, $\cF^\cl$ is same as $\cP$.  The path system $\cP$
    may be alternatively denoted in terms of $\cF$ and $\cl$ as
    $\cF^\cl$. In most scenarios in this paper, what is given are the
    pair $(\cF, \cl)$ and the target tree $T$; hence this notation
    will be used more often.}

%  \noindent
  If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$ is
  called a {\em path hypergraph} and $\cP$ is called {\em path
    representation} of $\cF$. If this isomorphism is $\phi: supp(\cF)
  \rightarrow V(T)$, then it is clear that there is an {\em induced
    path labeling} $\cl_\phi: \cF \rightarrow \cP$ to the set system;
  $\cl_\phi(S) = \set{y \mid y = \phi(x), x \in S}$ for all $S \in
  \cF$. Recall that $supp(\cP) = V(T)$.

%  \noindent
  \add{A graph} $G$ is a {\em path graph} if it is isomorphic to the
  intersection graph $\bI(\cP)$ of a path system $\cP$.  This
  isomorphism gives a bijection $\cl': V(G) \rightarrow
  \cP$. Moreover, for the purposes of this paper, we require that in a
  path labeling, $supp(\cP) = V(T)$.  If graph $G$ is also isomorphic
  to $\bI(\cF)$ for some hypergraph $\cF$, then clearly there is a
  bijection $\cl: \cF \rightarrow \cP$ such that $\cl(S) = \cl'(v_S)$
  where $v_S$ is the vertex corresponding to set $S$ in $\bI(\cF)$ for
  any $S \in \cF$. This bijection $\cl$ is called the {\em path
    labeling} of the hypergraph $\cF$ and the path system $\cP$ may be
  alternatively denoted as $\cF^\cl$.

%  \noindent
  A path labeling $(\cF, \cl)$ is defined to be {\em feasible} if $\cF
  \cong \cF^\cl$ and this hypergraph isomorphism $\phi: supp(\cF)
  \rightarrow supp(\cF^\cl)$ induces a path labeling $\cl_\phi: \cF
  \rightarrow \cF^\cl$ such that $\cl_\phi = \cl$.  \add{In this
    work}, we are given as input $\cF$ and a tree $T$, and the
  question is whether there is a path labeling $\cl$ to a set of paths
  in $T$.  We refer to such a solution path system by $\cF^\cl$.  A
  path labeling $(\cF, \cl)$ is defined to be {\em feasible} if
  % $\cF \cong \cF^\cl$ and this
  there is a hypergraph isomorphism $\phi: supp(\cF) \rightarrow
  supp(\cF^\cl)=V(T)$ induces a path labeling $\cl_\phi: \cF
  \rightarrow \cF^\cl$ such that $\cl_\phi = \cl$.
%\end{definition}

\subsection{Overlap Graphs and Marginal Hyperedges}
%\begin{definition}[Overlap Graphs and Marginal Hyperedges]
  An {\em overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is a graph
  such that its vertex set has a bijection to $\cF$ and there exists
  an edge between two of its vertices \iff their corresponding
  hyperedges overlap. Two hyperedges $S$ and $S'$ are said to {\em
    overlap}, denoted by $S \overlap S'$, if they have a non-empty
  intersection and neither is contained in the other; $S \overlap S'
  \text{ \iff } S \cap S' \ne \emptyset, S \nsubseteq S', S'
  \nsubseteq S$. Thus $\bO(\cF)$ is a spanning subgraph of $\bI(\cF)$
  and not necessarily connected. Each connected component of
  $\bO(\cF)$ is called an {\em overlap component}.

%  \noindent
  A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
  \overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
  \cite{kklv10}. Additionally, if $S$ is such that it is contained in
  no other marginal hyperedge in $\cF$, then it is called {\em
    super-marginal}.
%\end{definition}

\subsection{Miscellaneous}
%\begin{definition}[$k$-subdivided star] 
  A {\em star} graph is a
  complete bipartite graph $K_{1,p}$ which is clearly a tree and $p$
  is the number of leaves. The vertex with maximum degree is called
  the {\em center} of the star and the edges are called {\em rays} of
  the star. A {\em $k$-subdivided star} is a star with all its rays
  subdivided exactly $k$ times. The definition of a {\em ray of a
    $k$-subdivided star} is extended to the path from the center to a
  leaf. It is clear that all rays are of length $k+2$.
%\end{definition}

@@

%\begin{definition}[Partially ordered sets]

  Let $X$ be a partially ordered set with $\preccurlyeq$ being the
  partial order on $X$.  $X_m \in X$ is a maximal upper bound of $X$
  if $\nexists X_q \in X$ such that $X_m \preccurlyeq X_q$.  A maximal
  upper bound on $X$ is denoted by $mub(X)$.
  %represents an element in $X$ which is a maximal upper bound on $X$.
%\end{definition}

An {\em in-tree} is a directed rooted tree in which all edges are
directed toward to the root.

The set $I$ represents the index set $[m]$. If index $i$ is used
without further qualification, it is intended that $i \in I$.

@@

\tnote{\sc refer to Problem Defs here}
In this section we give an algorithmic characterization of the
feasibility of a tree path labeling.  
We focus on the question of generalizing the notion of an
  ICPIA cite-nsnrs09 to characterize feasible path assignments.  We
  show that for a given set system $\cF$, a tree $T$, and an
  assignment of paths from $T$ to the sets, there is a feasible
  bijection between $U$ and $V(T)$ if and only if all intersection
  cardinalities among any three sets (not necessarily distinct) is
  same as the intersection cardinality of the paths assigned to them
  and the input runs a filtering algorithm (described in this paper)
  successfully.  This characterization is proved constructively and it
  gives a natural data structure that stores all the relevant feasible
  bijections between $U$ and $V(T)$.  Further, the filtering algorithm
  is also an efficient algorithm to test if a tree path labeling to
  the set system is feasible.  This generalizes the result in
  cite-nsnrs09.

Consider a path labeling $(\cF,
\cl)$ on the given tree $T$. We call $(\cF, \cl)$ an {\em Intersection
  Cardinality Preserving Path Labeling (ICPPL)} if it has the
following properties.

\begin{enumerate}[{(\icpplpr\ }i)]
\item \label{pr:i} $|S| = |\cl(S)|$ for all $S \in \cF$
  % \vspace{\topshrink}
\item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$ for all
  distinct $S_1, S_2 \in \cF$
  % \vspace{\topshrink}
\item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
  \cap \cl(S_3)|$ for all distinct $S_1, S_2, S_3 \in \cF$
\end{enumerate}

In the remaining part of this section we show that $(\cF, \cl)$ is
feasible if and only if it is an ICPPL and
Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty path
hypergraph isomorphism
function. Algorithm~\ref{al:icppl-find-isomorph} recursively does two
levels of filtering of $(\cF, \cl)$ to make it simpler in terms of set
intersections while retaining the set of isomorphisms, if any, between
$\cF$ and $\cF^\cl$.
% One direction of this claim isclear: that if a path labeling is
% feasible, then all intersection cardinalities are preserved,
% i.e. the path labeling is an ICPPL. Algorithm~\ref{al:filterone}
% \annote{has no premature exit condition hence any input will go
%   through it}{Prove that the filtered sets has ICPPL \iff input PL
%   has ICPPL?}. Algorithm~\ref{al:filtertwo} has an exit condition at
% line~\ref{xempty}. It can be easily verified that $X$ cannot be
% empty if $\cl$ is a feasible path labeling. The reason is that a
% feasible path labeling has an associated bijection between
% $supp(\cF)$ and $V(T)$ \remove{i.e. $supp(\cF^{\cl})$} such that the
% sets map to paths, ``preserving'' the path labeling.  The rest of
% the section is devoted to constructively proving that it is
% sufficient for a path labeling to be an ICPPL and pass the two
% filtering algorithms.  To describe in brief, the constructive
% approaches refine an ICPPL iteratively, such that at the end of each
% iteration we have a ``filtered'' path labeling, and finally we have
% a path labeling that defines a family of bijections from $supp(\cF)$
% to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.
First, we present Algorithm~\ref{al:filterone} which describes {\tt
  filter\_1}, and prove its correctness.  This algorithm refines the
path labeling by processing pairs of paths in $\cF^\cl$ that share a
leaf until no two paths in the new path labeling share any leaf.

@@

    \IF{$(\cF_j, \cl_j)$ does not satisfy (\icpplpr~\ref{pr:iii}) of
      ICPPL}
    \label{ln:3waycheck}
    \STATE {\bf exit} \label{ln:exit1} \\
    \ENDIF

@@

\begin{lemma}
  \label{lem:feasible}
  In Algorithm~\ref{al:filterone}, if input $(\cF, \cl)$ is a feasible path
  assignment then at the end of $j$th iteration of the {\bf while}
  loop, $j \ge 0$, $(\cF_j, \cl_j)$ is a feasible path assignment.
\end{lemma}
\begin{proof}\thesisspacing%[Proof of Lemma~\ref{lem:feasible}]
  We will prove this by mathematical induction on the number of
  iterations. The base case $(\cF_0, \cl_0)$ is feasible since it is
  the input itself which is given to be feasible. Assume the lemma is
  true till $j-1$th iteration. \ie every hypergraph isomorphism $\phi:
  supp\left(\cF_{j-1}\right) \rightarrow V\left(T \right)$ that
  defines $(\cF, \cl)$'s feasibility, is such that the induced path
  labeling on $\cF_{j-1}$, say denoted by $\cl_{\phi[{\cF_{j-1}}]}$, is equal to
  $\cl_{j-1}$. We will prove that $\phi$ is also the bijection that
  makes $(\cF_j, \cl_j)$ feasible. Note that $supp(\cF_{j-1}) =
  supp(\cF_{j})$ since the new sets in $\cF_j$ are created from basic
  set operations to the sets in $\cF_{j-1}$ adding or removing no
  elements. For the same reason and $\phi$ being a bijection, it is
  clear that when applying the $\phi$-induced path labeling on
  $\cF_j$, $ \cl_{\phi[{\cF_{j}}]}(S_1 \setminus S_2) =
  \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
  \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Now observe that $ \cl_j(S_1
  \setminus S_2) = \cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2) =
  \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
  \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Thus the induced path labeling
  $\cl_{\phi[{\cF_{j}}]} = \cl_{j}$.  %\qed
\end{proof}

>>>>

A note on attribution -- the machinery for labeling of hypergraphs
has been adopted from \cite{kklv10} and adapted to the needs of our
work. Some classic graph definitions are from \cite{mcg04}.

\subsection{Set systems and Hypergraphs}
\label{sec:setsyshypergraph}

Consider a universe $U$ with $|U| = n$.  The set $\F \subseteq (2^{U}
\setminus \emptyset)$ is a {\em set system} with universe $U$.  The
{\em support of set system} $\F$ denoted by $\supt{\cF}$, is the union
of all the sets in $\F$.
\[supp(\F) = \bigcup_{S \in \F}S\]%
By convention, the set system $\cF$ ``covers'' its universe:
$supp(\cF) = U$. If the sets in $\cF$ are intervals or paths of a
tree, then it is called an interval system or a path system,
respectively.
\par
A set system $\cF$ can also be visualized as a {\em hypergraph}
$\cF_H$ whose vertex set is $U$ and hyperedges are the sets in
$\cF$. This is a known representation for interval systems in
literature \cite{bls99,kklv10}.  We extend this definition here to
path systems. Due to the equivalence of set system and hypergraph as
far as this thesis is concerned, we drop the subscript $_H$ in the
notation and refer to both the structures by $\cF$. Thus hypergraphs
and set systems are synonyms. So are hyperedges and sets.
\par
The notion of {\em hypergraph isomorphism} is central to the work of
this thesis. This is defined by Definition~\ref{def:hisomorph}.

\begin{definition}[Hypergraph isomorphism]
  \label{def:hisomorph}
  Let $\cF'$, $\cF''$ be two hypergraphs.  They are said to be {\em
    isomorphic} to each other, denoted by $\cF' \cong \cF''$, \iff
  there exists a bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$
  such that for all sets $A \subseteq supp(\cF')$, $A$ is a hyperedge
  in $\cF'$ \iff $B$ is a hyperedge in $\cF''$ where $B = \{\phi(x)
  \mid x \in A\}$ (this is also written as $B=\phi(A)$).
\end{definition}

Figure~\ref{fig:hypergraphs} shows an example of hypergraphs and
hypergraph isomorphism.

Recall intersection graph and path graph from
Definition~\ref{def:intersectiongraph} and \ref{def:pathgraph}. The
{\em intersection graph} $\bI(\cF)$ of a hypergraph or set system
$\cF$ is a graph such that its vertex set has a bijection to $\cF$ and
there exists an edge between two vertices \iff their corresponding
hyperedges have a non-empty intersection. A graph $G$ is a path graph
if it is isomorphic to the intersection graph of a path system.

\begin{figure}[htb]
  \centering

  \begin{tabular}[h]{l}

    \begin{tabular}[h]{ll}
      \begin{tabular}[t]{l}
        $\cF_1:$ \\
        \hypergraphi%
      \end{tabular} 
      &     
      \begin{tabular}[t]{l}
        $\cF_2:$\\
        \hypergraphii %        
      \end{tabular}
    \end{tabular}
    \\
    \begin{tabular}[h]{ll}
      \begin{tabular}[t]{l}
        \\ \\
        $\cP$:\\
        \pathhypergraphi %       
      \end{tabular}    
      &     
      \begin{tabular}[t]{l}
        $\cF_3:$\\
        \hypergraphiii %        
      \end{tabular}
    \end{tabular}
    \\  
    \begin{tabular}[h]{l}
      $\cI$:\\
      \intervalhypergraphi %
    \end{tabular}
  \end{tabular}

  \caption[\figtabsize Hypergraphs and set systems]{\figtabsize
    Hypergraphs and set systems. $\cF_1$ and $\cF_2$ are isomorphic
    hypergraphs. The subscript of each element is such that if $\phi$
    is the hypergraph isomorphism, $\phi(x_i)=y_i$. $\cI$ and $\cP$
    are an interval system and a path system respectively which are both
    isomorphic to $\cF_1$ (and by transitivity, isomorphic to $\cF_2$). Here too the
    subscript of elements indicate the isomorphism bijection. $\cF_3$
    is a hypergraph that is not isomorphic to any of the other
    hypergraphs -- moreover, it is one that cannot have a feasible
    path labeling to any tree since it does not have the Helly's
    property anymore due to the addition of hyperedge $\{x_1,
    x_4\}$. }
  \label{fig:hypergraphs}
\end{figure}

\subsection{Path Labeling and Path Hypergraphs}
\label{sec:pathhypergraph}
The graph $T$ represents a {\em target tree} with same number of
vertices as elements in $U$; $|V(T)|=|U|=n$.  A {\em path system}
$\cP$ is a set system of paths from $T$. 
\[\cP \subseteq \{P \mid P \subseteq V, \text{ } T[P] \text{ is a
  path} \}\]

A special mapping of paths from target tree $T$ to hyperedges in $\cF$
is called a {\em path labeling}. Definition~\ref{def:pathlabel} gives
details. Note that the term ``tree path labeling'' is synonymous
with ``path labeling'' in this document since we only consider trees and
not general graphs as source of paths.

\begin{definition}[Path labeling of a hypergraph]
  \label{def:pathlabel}
  Consider a hypergraph $\cF$ % with universe $U$
  and a path system $\cP$ from tree $T$ such that $\bI(\cF) \cong
  \bI(\cP)$.  Then the associated bijection $\cl: \cF
  \rightarrow \cP$ due to this isomorphism is called a {\em path
    labeling} of $\cF$ from $T$. The path system $\cP$ is
  alternatively denoted as $\cF^\cl$. The path labeling may also be
  denoted by ordered pairs $(\cF,\cl)$ or $(\cF,\cl,T)$.
\end{definition}
% If the intersection graphs of $\cF$ and $\cP$ are isomorphic,
% $\bI(\cF) \cong \bI(\cP)$, then the associated bijection $\cl: \cF
% \rightarrow \cP$ due to this isomorphism is called a {\em path
%   labeling} of the hypergraph $\cF$.
To elaborate on the phrase ``associated bijection'' in
Definition~\ref{def:pathlabel}, let $\cg: V(\bI(\cF)) \rightarrow
V(\bI(\cP))$ be the given intersection graph isomorphism, where
$V(\bI(\cF)) = \{ v_S \mid S \in \cF\}$ and $V(\bI(\cP)) = \{v_P \mid
P \in \cP\}$. Then the associated bijection $\cl: \cF \rightarrow \cP$
is defined as follows. For all $S' \in \cF$ and $P' \in \cP$, $\cl(S')
= P'$ \iff $\cg(v_{S'}) = v_{P'}$.  Thus, it can be seen that a path
labeling preserves the non-empty intersection property between every
pair of hyperedges in $\cF$.

At this point, it is worth comparing Definition~\ref{def:pathlabel}
with Definition~\ref{def:hisomorph} seen in the previous section.  It
may be noted that there are two kinds of isomorphisms here. One is the
isomorphism of intersection graphs of $\cF$ and $\cP$, i.e. $\bI(\cF)$
and $\bI(\cP)$ and second is the isomorphism between the hypergraphs
$\cF$ and $\cP$ themselves. It is easy to verify that given a path
labeling of hypergraph $\cF$, there need not be a hypergraph
isomorphism involved. In fact, checking if an induced hypergraph
isomorphism results from a given path labeling is the problem of
\CFTPL.  However, a hypergraph isomorphism always induces a path
labeling as shown by the following Definition~\ref{def:inducedpl}.

\begin{definition}[Hypergraph isomorphism induced path labeling]
  \label{def:inducedpl}
  Let $\cF$ be a hypergraph and $\cP$ be a path system \stt $\cF \cong
  \cP$. Let  $\phi: \supt{\cF} \rightarrow \supt{\cP}$ be this hypergraph
  isomorphism function. % described in
  % Definition~\ref{def:pathhypergraph}.
  Then there is an {\em induced path labeling} % denoted by $\cl_\phi:
%   \cF \rightarrow \cP$ 
  denoted by $\cp_\phi: \cF \rightarrow \cP$
  of the hypergraph \stt % $\cl_\phi(S) = \phi(S)$
  $\cp_\phi(S) = \phi(S)$
  (see Definition~\ref{def:hisomorph}) for all $S \in \cF$.  
\end{definition}

Combining the ideas of path labeling, hypergraph isomorphism and its
induced path labeling we now define {\em feasiblity} of a path
labeling in Definition~\ref{def:feasibletpl}.

\begin{definition}[Feasible path labeling]
  \label{def:feasibletpl}
  Let $\cF$ be a hypergraph and $(\cF, \cl)$ be a path
  labeling. $(\cF, \cl)$ is called a {\em feasible path labeling} if
  the following conditions hold.
  \begin{itemize}
  \item $\cF \cong \cF^\cl$
  \item if the above hypergraph isomorphism is $\phi: supp(\cF)
    \rightarrow supp(\cF^\cl)$, then $\cp_\phi = \cl$
  \end{itemize}

  This isomorphism $\phi$ is called the \textbf{feasibility hypergraph
  isomorphism} or feasibility isomorphism.
\end{definition}

One of the problems solved in this thesis \FTPL characterizes
feasibile tree path labeling.

Now we will see the definition of a {\em path hypergraph} which is a
simple idea extended from interval hypergraphs in \cite{kklv10}.

\begin{definition}[Path hypergraph]
  \label{def:pathhypergraph}
  Let $\cF$ be a set system. $\cF$ is called a {\em path hypergraph}
  if there exists a target tree $T$ and a path system $\cP$ from $T$
  \stt $\cF \cong \cP$. If $\cF$ is thus a path hypergraph, $\cP$ is
  called {\em path representation} of $\cF$.
\end{definition}

In other words, if there exists a feasible tree path labeling for a
given hypergraph, it is a path hypergraph.

%% XXXXXX
% The subset of the path labeling that
%   maps elements of the universe to paths that contain only a leaf the
%   target tree is  known as a {\em leaf labeling}.
%   \ie if the target tree has the set of $k$ leaves $L = \{v_i, i \in
%   [k]\}$, then 
% $\cl': \rightarrow U' \rightarrow L$ is a leaf labeling where $U'
% \subseteq U$, $|U'| = k$.
%   $\cL_0 = \{ (x_i,v_i) \mid x_i \in
%   U, v_i \in T, x_i \ne x_j, v_i \ne v_j, i \ne j, i,j \in [k] \}$
%   where $k$ is the number of leaves in $T$

\subsection{Overlap Graphs and Marginal Hyperedges}
\label{sec:overlapmarginal}
Recall the notion of {\em overlap} from
Definition~\ref{def:setsystem}.  Two hyperedges $S$ and $S'$ are said
to overlap, denoted by $S \overlap S'$, if they have a non-empty
intersection and neither is contained in the other.

An {\em overlap graph}  $\bO(\cF)$ of a hypergraph $\cF$ is a graph
such that its vertex set has a bijection to $\cF$ and there exists an
edge between two of its vertices \iff their corresponding hyperedges
overlap.  Thus $\bO(\cF)$ is a spanning subgraph of $\bI(\cF)$ and not
necessarily connected. Each connected component of $\bO(\cF)$ is
called an {\em overlap component}.

A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
\overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
\cite{kklv10} (see Definition\ref{def:poset}). Additionally, if $S$ is
such that it is contained in no other marginal hyperedge in $\cF$,
then it is called {\em super-marginal}.

\subsection{Miscellaneous}
\label{sec:miscprelim}
A {\em star} graph is a complete bipartite graph $K_{1,\,p}$ (which is
clearly a tree with $p$ leaves). The vertex with maximum degree in a
star is called its {\em center} and the edges are called {\em
  rays}. This thesis is interested in a star-like graph which is
defined in Definition~\ref{def:ksubstar}

\begin{definition}[$k$-subdivided star]
  \label{def:ksubstar}
  A {\em $k$-subdivided star} is a star with all its rays subdivided
  exactly $k$ times. A {\em ray} of a $k$-subdivided star is the path
  from the center vertex to a leaf. It is clear that all rays of a
  $k$-subdivided star are of length $k+2$.
\end{definition}

@@

\begin{definition}[In-tree]
An {\em in-tree} is a directed rooted tree in which all edges are
directed toward to the root.  
\end{definition}
In-trees are used in Section~\ref{sec:decompos} which describes our
solution to \CFTPL problem.

Finally, note that as a convention in this thesis, the set $I$
represents the index set $[m]$ or $[n]$. If index $i$ is used without further
qualification, it is intended that $i \in I$.

@@

In this section we discuss the \FTPL problem introduced in
Section~\ref{sec:myresearchintro} and give a set cardinality based
characterization for it.

For this characterization, we generalize the notion of the ICPIA
\cite{nsnrs09}.  We show that for a given hypergraph $\cF$ with
universe $U$, a target tree $T$, and path labeling from $T$ to $\cF$, there
is a feasible bijection between $U$ and $V(T)$ \iff all intersection
cardinalities among any three sets (not necessarily distinct) is same
as the intersection cardinality of the paths assigned to them.  This
characterization is proved constructively and the algorithm outputs a
feasible bijection between $U$ and $V(T)$.

We now introduce a special kind of path labeling called {\em
  Intersection Cardinality Preserving Path Labeling (ICPPL)} in
Definition~\ref{def:icppl}.

\begin{definition}
  \label{def:icppl}
  Consider a path labeling $(\cF, \cl)$ on the given tree $T$. We call
  $(\cF, \cl)$ an {\em Intersection Cardinality Preserving Path
    Labeling (ICPPL)} if it has the following properties.
  \par
  \parbox{\linewidth}{%
    \hangindent \defindent%
    \begin{enumerate}[i. ] %{(\icpplpr\ }
      \singlespacing
    \item \label{pr:i} $|S| = |\cl(S)|$ for all $S \in \cF$
    \item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$ for
    all distinct $S_1, S_2 \in \cF$
    \item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap
    \cl(S_2) \cap \cl(S_3)|$ for all distinct $S_1, S_2, S_3 \in \cF$
    \end{enumerate}
  }
\end{definition}

In the remaining part of this section we show that $(\cF, \cl)$ is
feasible if and only if it is an ICPPL%  and
% Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty path
% hypergraph isomorphism
% function
.  Algorithm~\ref{al:icppl-find-isomorph} takes $(\cF, \cl)$ as input
and computes the feasibility bijection if it is feasible or else
reports failure. This algorithm recursively does two levels of
filtering of $(\cF, \cl)$ to make it simpler in terms of set
intersections while retaining a non-empty subset of isomorphisms, if
any, between $\cF$ and $\cF^\cl$. These two filters are {\tt
  filter\_1} in Algorithm~\ref{al:filterone} and {\tt filter\_2} in
Algorithm~\ref{al:filtertwo}.
% One direction of this claim isclear: that if a path labeling is
% feasible, then all intersection cardinalities are preserved,
% i.e. the path labeling is an ICPPL. Algorithm~\ref{al:filterone}
% \annote{has no premature exit condition hence any input will go
%   through it}{Prove that the filtered sets has ICPPL \iff input PL
%   has ICPPL?}. Algorithm~\ref{al:filtertwo} has an exit condition at
% line~\ref{xempty}. It can be easily verified that $X$ cannot be
% empty if $\cl$ is a feasible path labeling. The reason is that a
% feasible path labeling has an associated bijection between
% $supp(\cF)$ and $V(T)$ \remove{i.e. $supp(\cF^{\cl})$} such that the
% sets map to paths, ``preserving'' the path labeling.  The rest of
% the section is devoted to constructively proving that it is
% sufficient for a path labeling to be an ICPPL and pass the two
% filtering algorithms.  To describe in brief, the constructive
% approaches refine an ICPPL iteratively, such that at the end of each
% iteration we have a ``filtered'' path labeling, and finally we have
% a path labeling that defines a family of bijections from $supp(\cF)$
% to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.

First, we present Algorithm~\ref{al:filterone} which describes {\tt
  filter\_1}, and prove its correctness.  This algorithm refines the
path labeling by processing pairs of paths in $\cF^\cl$ that share a
leaf until no two paths in the new path labeling share any leaf.

@@

%     \IF{$(\cF_j, \cl_j)$ does not satisfy (\icpplpr~\ref{pr:iii}) of
%       ICPPL}
%     \label{ln:3waycheck}
%     \STATE {\bf exit} \label{ln:exit1} \\
%     \ENDIF

@@

Now we will prove the effectiveness of Algorithm~\ref{al:filterone}.
Lemma~\ref{lem:feasible} proves that {\tt filter\_1} does not destroy
the feasibility of the input, if it is indeed feasible. Following
this, Lemma~\ref{lem:invar1} and Lemma~\ref{lem:noexit1} show that if
the input is an ICPPL the result of {\tt filter\_1} is also an ICPPL
and they both have the same set of feasibility hypergraph
isomorphisms.

\begin{lemma}
  \label{lem:feasible}
  In Algorithm~\ref{al:filterone}, if input $(\cF, \cl)$ is a feasible path
  labeling then at the end of $j$th iteration of the {\bf while}
  loop, $j \ge 0$, $(\cF_j, \cl_j)$ is a feasible path labeling.
\end{lemma}
\begin{proof}\thesisspacing%[Proof of Lemma~\ref{lem:feasible}]
  We will prove this by mathematical induction on the number of
  iterations. The base case $(\cF_0, \cl_0)$ is feasible since it is
  the input itself which is given to be feasible. Assume the lemma is
  true till $j-1$th iteration \ie every hypergraph isomorphism $\phi:
  supp\left(\cF_{j-1}\right) \rightarrow V\left(T \right)$ that
  defines $(\cF, \cl)$'s feasibility, is such that the induced path
  labeling on $\cF_{j-1}$, say denoted by $\cp_{\phi[{\cF_{j-1}}]}$,
  is equal to $\cl_{j-1}$.
% Note that any hypergraph isomorphism of
% $\cF$ (say, one that defines the feasibility of $(\cF,\cl)$) can be
% ``projected'' to $\cF_{j-1}$ since their supports are equal. If $\phi$
% is a hypergraph isomorphism of to $(\cF, \cl)$, then 
  We will prove that $\phi$ is also the bijection that makes $(\cF_j,
  \cl_j)$ feasible. Note that $supp(\cF_{j-1}) = supp(\cF_{j})$ since
  the new sets in $\cF_j$ are created from basic set operations to the
  sets in $\cF_{j-1}$ adding or removing no elements. For the same
  reason and $\phi$ being a bijection, it is clear that when applying
  the $\phi$-induced path labeling on $\cF_j$, $
  \cp_{\phi[{\cF_{j}}]}(S_1 \setminus S_2) =
  \cp_{\phi[{\cF_{j-1}}]}(S_1) \setminus
  \cp_{\phi[{\cF_{j-1}}]}(S_2)$. Now observe that $ \cl_j(S_1
  \setminus S_2) = \cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2) =
  \cp_{\phi[{\cF_{j-1}}]}(S_1) \setminus
  \cp_{\phi[{\cF_{j-1}}]}(S_2)$. Thus the induced path labeling
  $\cp_{\phi[{\cF_{j}}]} = \cl_{j}$.  %\qed
\end{proof}
